---
title: "Instacart Market Basket Analysis Data Viz"
author: "Group 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(data.table)
library(knitr)
library(tidyverse)
library(modelsummary)
library(gridExtra)

setwd("/Users/jinglyng/Documents/MSBA/Semester 2/CS5344/Group Project/instacart-market-basket-analysis")
aisle <- fread("aisles.csv")
dept <- fread("departments.csv")
order_pdts_prior <- fread("order_products__prior.csv")
order_pdts_train <- fread("order_products__train.csv")
orders <- fread("orders.csv")
products <- fread("products.csv")
```

### Understanding the data
**Aisles**
There are 134 aisles in the dataset
```{r}
str(aisle)
paste(sort(head(aisle$aisle)), collapse=', ')
```
**Departments**
There are 21 departments in the dataset
```{r}
str(dept)
paste(sort(head(dept$department)), collapse=', ')
```
**Products**
There are 49,688 products in the catalogue.
Products are sorted into their respective aisles and departments. 
```{r}
str(products)
kable(head(products))
```

**Orders**
There are ~3.4 million observations in the order dataset. The observations are split into "prior", "train", and "test" sets. 

```{r}
str(orders)
unique(orders$eval_set)
```
Analysing the orders of user number 1, we can see that the orders dataset shows that the user has 10 prior orders and 1 train order. Looking at the other users, it seems like the latest order of the users will either be the train or test set. 
```{r}
orders[user_id == 1]
```

**Order_Product**
There are two datasets that combine order and product information. 
- `orders_products_prior`: data on all prior orders
- `orders_products_train`: data on the train sets (last order)

This dataset breaks down the orders into each individual item. 
`add_to_cart_order` tells us which order the product was added to card, 
`reordered` tells us which products have been reordered by the user before.

```{r}
str(order_pdts_prior)
head(order_pdts_prior)
```

**Prior and Train Set**
```{r}
# Number of Users in Each Set 
orders[eval_set == "prior", length(unique(user_id))] # 206,209
orders[eval_set == "train", length(unique(user_id))] # 131,209
## Presumably, the missing users are in the test set that is not given to us

# Number of Orders in Each Set
orders[eval_set == "prior", length(unique(order_id))] # 3,214,874
orders[eval_set == "train", length(unique(order_id))] # 131,209

# Number of Obs for Orders_Products in Each Set
order_pdts_prior[, .N] # 32,434,489
order_pdts_train[, .N] # 1,384,617
```

Since we will not be training a ML model, we will just merge all the data together for EDA.

```{r}
dt <- rbind(order_pdts_prior, order_pdts_train)
dt <- merge(dt, orders, by = "order_id")
dt <- merge(dt, products, by = "product_id")
dt <- merge(dt, aisle, by = "aisle_id")
dt <- merge(dt, dept, by = "department_id")

dt <- dt[, c(7, 4, 3, 5, 6, 9, 10, 11, 12, 13, 1, 15, 2, 14, 8)]
setorder(dt, user_id, order_number, order_id,  add_to_cart_order)
```

### Subsetting the full dataset for apiori mechanism
Since we are finding apriori rules, we will have to get all the products within an order. So we will randomly select 10% of all the orders.
```{r apriori_subset}
set.seed(123)
n <- round(0.1 * orders[, .N])
orders_sample <- sample_n(orders, n)
dt_apriori <- dt[order_id %in% orders_sample$order_id]
```

We'll need to check that the sampled data is comparable to the actual dataset

```{r}
# Checking the top 10 products
tmp <- unique(dt_apriori[, total_ordered := length(order_id), product_name][, percentage := total_ordered/length(unique(order_id))][, .(product_name, total_ordered, percentage)])


plot1 <- ggplot(data = top_n(tmp, 15), aes(x = reorder(product_name, percentage), y = percentage)) +
  geom_col() +
  labs(y = "",
       title = "Top 15 Products Apriori Sample") +
    theme (
    axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
    axis.title.x = element_blank()) 

tmp2 <- unique(dt[, total_ordered := length(order_id), product_name][, percentage := total_ordered/length(unique(order_id))][, .(product_name, total_ordered, percentage)])

plot2 <- ggplot(data = top_n(tmp2, 15), aes(x = reorder(product_name, percentage), y = percentage)) +
  geom_col() +
  labs(y = "",
       title = "Top 15 Products (Full Dataset)") +
    theme (
    axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
    axis.title.x = element_blank()) 

grid.arrange(plot1, plot2, ncol=2)

# Checking number of products in each aisle and dept
aisle_apr <- unique(dt_apriori[, total_ordered := length(order_id), aisle][, percentage := total_ordered/length(unique(order_id))][, .(aisle, total_ordered, percentage)])

plot3 <- ggplot(data = top_n(aisle_apr, 15), aes(x = reorder(aisle, percentage), y = percentage)) +
  geom_col() +
  labs(y = "",
       title = "Top 15 Aisles Apriori Sample") +
    theme (
    axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
    axis.title.x = element_blank()) 

aisle_dt <- unique(dt[, total_ordered := length(order_id), aisle][, percentage := total_ordered/length(unique(order_id))][, .(aisle, total_ordered, percentage)])

plot4 <- ggplot(data = top_n(aisle_dt, 15), aes(x = reorder(aisle, percentage), y = percentage)) +
  geom_col() +
  labs(y = "",
       title = "Top 15 Aisles") +
    theme (
    axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
    axis.title.x = element_blank()) 

grid.arrange(plot3, plot4, ncol=2)

dept_apr <- unique(dt_apriori[, total_ordered := length(order_id), department][, percentage := total_ordered/length(unique(order_id))][, .(department, total_ordered, percentage)])

dept_dt <- unique(dt[, total_ordered := length(order_id), department][, percentage := total_ordered/length(unique(order_id))][, .(department, total_ordered, percentage)])
setnames(dept_dt, "percentage", "percentage_all")

dept <- merge(dept_apr, dept_dt, by = "department")
dept <- dept[, .(department, percentage, percentage_all)]

setorder(dept, percentage)
print(dept)

# fwrite(dt_apriori, "apriori_subset.csv")
```

## Exploratory Data Analysis
Interesting questions to explore:\

**Products**
- What are the top products ordered? What about the top products reordered?
- What are the top products ordered everyday? What about time of the day?
- Are there products that have never been ordered? Are there products that have never been reordered? 
- How many products are reordered? 

```{r}
dt%>%
  group_by(order_dow, product_name) %>%
    summarize(n=n()) %>%
  mutate(percentage=n/sum(n)) %>%
  top_n(10, wt=n) %>%
  ggplot (aes(x=as.factor(order_dow), y=percentage, fill=product_name)) + 
  geom_col() + 
  labs(x = "Days of the Week", 
       y = "Proportion of Orders in a Day",
       title = "Daily Top 10 Products ordered") +
  theme(legend.position="bottom",legend.direction="horizontal")
```

**Orders**
- How many orders did each user make? 
- How long did it take before the user made the next order?  
- 

```{r}
tmp = dt %>% group_by(user_id) %>% summarize(n_orders = max(order_number))
tmp %>% ggplot(aes(x=as.factor(n_orders))) + 
  geom_bar(colour = 'darkblue', fill='darkblue') +
  labs(x = "Number of Orders Made by Users", 
       y = "Count of Users",
       title = "How many orders did each user make?") +
  theme(
      axis.text.x  = element_text (size = 6.0, angle = (90), hjust = 1, vjust = 0.5)
    )

tmp = dt %>% 
  filter(order_number>1) %>% 
  group_by(order_id) %>% 
  summarize(n_orders = max(days_since_prior_order)) 

tmp %>% ggplot(aes(x=as.factor(n_orders))) + 
  geom_bar(colour = 'darkgreen', fill='darkgreen') + 
    labs(
      x = "Days Since Previous Order", 
      y = "Count of Orders",
      title = "How long did it take before the user made the next order?")
```
## Cart Size
```{r}
tmp = dt %>%
  group_by(order_id)  %>%
  summarize( basket_size=n(), 
             reordered_items = sum(reordered)) %>%
  group_by(basket_size) %>%
  summarize(n=n(), avg_reordered_items =mean(reordered_items)) %>%
  arrange(basket_size)
  
tmp %>% ggplot(aes(x=as.factor(basket_size))) +
    geom_col(aes(y=n), colour = "darkred", fill="darkred") +
    ylab('Count of Orders') +
    xlab('Number of Items in Basket') +
    ggtitle('Distribution of Basket Size') +
    theme(
      axis.text.x  = element_text (size = 6.0, angle = (90), hjust = 1, vjust = 0.5)
    )
```
